---
title: "NHST PRL analysis_26.05.25"
format: html
editor: visual
---

# Comparing frequentist and Bayesian approaches

Hypothetically, modelling the same data, an **NHST** might suggest a statistically significant effect (e.g., *p* \< .05), while a similar **Bayesian** model could suggest high uncertainty and no clear evidence of an effect (e.g., credible interval overlaps 0). This might happen when looking for a small effect in a large sample with minimal variance.

Here, we check to see if NHST might offer significant results. We do this to emphasize the utility of Bayesian modelling approaches when searching for small effects in heterogenious populations (e.g., autism).

# Data preparation

## Load data

Load in data:

```{r}
#Read in saved data file:
data <- read.csv("Data.csv")

#Prepare and standardise variables:
std_data <- data %>%
  mutate(
    EDAQ = standardize(EDAQ),
    MASQ = standardize(MASQ),
    RLLR = standardize(RLLR),
    perseveration = standardize(perseveration),
    regression = standardize(regression),
    loseShift = standardize(loseShift)
  ) %>%
  select(subject, EDAQ, MASQ, RLLR, loseShift, perseveration, regression, neurotype) %>%
  distinct()
```

# NHST analysis

Run NHST regression models for anxiety (MASQ), PDA behaviour (EDA-QA), and autism (neurotype) predicting lose-shift behaviour:

```{r}
#Fit the model where lose-shift behaviours are predicted by MASQ scores by neurotype:
nhstLSM <- lm(
  loseShift ~ MASQ*neurotype,
  data = std_data
  )

#Fit the model where lose-shift behaviours are predicted by a MASQ*neurotype and EDA-QA*neurotype:
nhstLSME <- lm(
  loseShift ~ MASQ*neurotype + EDAQ*neurotype,
  data = std_data
  )

#View output:
summary(nhstLSM)
summary(nhstLSME)
```

Run NHST regression models for anxiety (MASQ), PDA behaviour (EDA-QA), and autism (neurotype) predicting perseveration:

```{r}
#Fit the model where perseverative errors are predicted by MASQ scores by neurotype:
nhstPM <- lm(
  perseveration ~ MASQ*neurotype,
  data = std_data)

#Fit the model where perseverative errors are predicted by a MASQ*neurotype and EDA-QA*neurotype:
nhstPME <- lm(
  perseveration ~ MASQ*neurotype + EDAQ*neurotype,
  data = std_data )

#View output:
summary(nhstPM)
summary(nhstPME)
```

Run NHST regression models for anxiety (MASQ), PDA behaviour (EDA-QA), and autism (neurotype) predicting regression:

```{r}
#Fit the model where regressive errors are predicted by MASQ scores by neurotype:
nhstReM <- lm(
  regression ~ MASQ*neurotype,
  data = std_data  )

#Fit the model where regressive errors are predicted by a MASQ*neurotype and EDA-QA*neurotype:
nhstReME <- lm(
  regression ~ MASQ*neurotype + EDAQ*neurotype,
  data = std_data  )

#View output:
summary(nhstReM)
summary(nhstReME)
```

Run NHST regression models for anxiety (MASQ), PDA behaviour (EDA-QA), and autism (neurotype) predicting RLLR:

```{r}
#Fit the model where relative log learning rate (RLLR) are predicted by MASQ scores by neurotype:
nhstRM <- lm(
  RLLR ~ MASQ*neurotype,
  data = std_data)

#Fit the model where relative log learning rate (RLLR) are predicted by MASQ and EDA-QA scores by neurotype:
nhstRME <- lm(
  RLLR ~ MASQ*neurotype + EDAQ*neurotype,
  data = std_data  )

#View output:
summary(nhstRM)
summary(nhstRME)
```

# Conclusion

In most cases, the above frequentist results are far from significant. However, the effect of EDA-QA on RLLR is significant. Posterior estimates from the Bayesian version of this regression model are broad and overlap 0, suggesting uncertainty in this effect; the effect is not clearly distinguishable from no effect.

```{r}
#Load the model output from a file:
modRME <- readRDS("modRME_output.rds")

#Return model output:
summary(modRME)

#Read in posterior samples .csv to data frame:
modRME_posterior <- read.csv("modRME_posterior.csv")


#Filter out 'lprior' and 'lp__':
modRME_posterior <- modRME_posterior[, c("b_Intercept", "b_MASQ", "b_neurotypeNT","b_EDAQ", "b_MASQ.neurotypeNT", "b_neurotypeNT.EDAQ", "sigma")]

#Convert the data frame to long format:
modRME_posterior_long <- pivot_longer(modRME_posterior, 
                                        cols = everything(), 
                                        names_to = "Parameter", 
                                        values_to = "Estimate")

#Rename parameters to 'Intercept', 'MASQ', 'EDAQ', and 'sigma':
modRME_posterior_long$Parameter <- recode(modRME_posterior_long$Parameter,
                                            "b_Intercept" = "Intercept",
                                            "b_MASQ" = "MASQ", 
                                            "b_neurotypeNT" = "neurotypeNT",
                                            "b_EDAQ" = "EDAQ", 
                                            "b_MASQ.neurotypeNT" = "MASQ:neurotypeNT", 
                                            "b_neurotypeNT.EDAQ" = "neurotypeNT:EDAQ", 
                                            "sigma" = "Sigma")

#Extract group information from parameter names:
modRME_posterior_long$Group <- sub(":.*", "", modRME_posterior_long$Parameter)

#Reorder the parameters and flip the order:
modRME_posterior_long$Parameter <- factor(modRME_posterior_long$Parameter, 
                                            levels = rev(c("Intercept","MASQ","neurotypeNT","EDAQ","MASQ:neurotypeNT","neurotypeNT:EDAQ","Sigma")))

#Plot marginal posterior distributions with specified colors:
plot_title <- ggtitle("Posterior distributions",
                      subtitle = "with means and 95% HDI")

#Plot:
ggplot(modRME_posterior_long, aes(x = Estimate, y = Parameter, fill = neurotype)) +
  stat_halfeye(point_interval = mean_hdi, .width = c(0.95), size = 0.5, , height = 3, fill = "#9ED4D1") +  # Adjusting the size of the points
  #scale_fill_manual(values = colors) +
  labs(x = "Parameter estimates", y = "Parameters", fill = "Neurotype") +
  plot_title +
  theme_minimal() +  # Using theme_minimal() instead of theme_clean()
  theme(legend.position = "right",
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank(),  # Remove minor grid lines
        axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    plot.title = element_text(size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)) +  # Increase font size of axis titles) +  # Add a single line at 0 on the x-axis
  geom_vline(xintercept = 0, linetype = "dashed")  # Add a dashed line at x = 0
ggsave( 'RLLR_dist_EDAQ.png', plot = last_plot(), dpi = 300)
```

In this instance, leveraging a Bayesian approach to data modelling has utility over frequentist statistics, allowing for the quantification of uncertainty in parameter estimates (i.e., credible intervals); while frequentist statistics might indicate a significant effect, Bayesian analysis suggests we cannot be sure of the utility of this effect.
